### 2025-06-06 sec/related-work.tex

\section{Related Work}
\label{sec:related}

\todo{flesh out}

\begin{itemize}
\item Transparent visualisations \cite{perera22,bond25,psallidas18smoke}
\end{itemize}