### 2025-08-21 sec/related-work.tex
Import related-work changes from negative-cases.

\section{Related Work}
\label{sec:related}

\paragraph{Interpretable Natural Language Processing}

\begin{itemize}
    \item Transparent visualisations \cite{perera22,bond25,psallidas18smoke}
    \item Recent advances in interpretable language understanding, highlighting the shift from fine-tuning to prompt-based learning in NLP and presenting research on improving the transparency of neural models through interpretable classifiers, uncertainty analysis, and explainable systems. \cite{Yulan_2023}
    \item Quantity extraction in scientific text \cite{bolucu-etal-2023-investigating}
    \item Scientific fact checking \cite{abu-ahmad-etal-2025-climatecheck-shared}
\end{itemize}