### 2025-08-22 sec/related-work.tex
Bib fixes.

\section{Related Work}
\label{sec:related}

\paragraph{Interpretable NLP.}

Recent advances in interpretable language understanding, highlighting the shift from fine-tuning to
prompt-based learning in NLP and presenting research on improving the transparency of neural models through
interpretable classifiers, uncertainty analysis, and explainable systems \cite{yulan23}.

\paragraph{AI interpretation of scientific text.}

For example quantity extraction~\cite{bolucu23}, scientific fact-checking
\cite{abu-ahmad25}.

\paragraph{Linguistic competences of LLMs.}

\paragraph{Transparent programming languages.}

\cite{atkey25,perera22,bond25,psallidas18smoke}