### 2025-09-13 paper/tex/authors.tex
New authors.

\author{
   Alfonso Piscitelli \\
   University of Salerno \\
   Fisciano, Italy \\
   \texttt{apiscitelli@unisa.it} \\
   \And
   Joe Bond \\
   University of Bristol \\
   Bristol, UK \\
   \texttt{j.bond@bristol.ac.uk} \\
   \And
   Cristina David \\
   University of Bristol \\
   Bristol, UK \\
   \texttt{cristina.david@bristol.ac.uk} \\
   \And
   Ali Mohammed \\
   Manchester Metropolitan University \\
   Manchester, UK \\
   \texttt{22465707@stu.mmu.ac.uk} \\
   \And
   Jacob Pake \\
   University of Kent \\
   Canterbury, UK \\
   \texttt{jwp24@kent.ac.uk} \\
   \And
   Roly Perera \\
   University of Cambridge \\
   Cambridge, UK \\
   \texttt{roly.perera@cl.cam.ac.uk} \\
   \And
   Jessy Sodimu \\
   University of Bath \\
   Bath, UK \\
   \texttt{js3720@bath.ac.uk} \\
   \And
   Chenyiqiu Zheng \\
   University College London \\
   London, UK \\
   \texttt{chenyiqiu.zheng.23@ucl.ac.uk} \\
}

% ORCID IDs (not shown in document)
% Alfonso Piscitelli: 0000-0002-3567-1504
% Joe Bond: 0009-0008-5995-6994
% Cristina David: 0000-0002-9106-934X
% Roly Perera: 0000-0001-9249-9862