### 2025-09-18 paper/abstract.tex
Abstract tweaks.

\begin{abstract}
We introduce the idea of \emph{transparent documents}, web-based data-driven scholarly articles which allow
readers to explore the relationship to the underlying data by hovering over fragments of text. We present an
agent-based LLM framework for authoring transparent documents, building on recent developments in data
provenance for general-purpose programming languages. Our implementation uses Fluid, an open source functional
programming language with a provenance-tracking runtime, as a target platform. Our tool consists of two LLM
agents which support a human author during the creation of a transparent document. A \SuggestionAgent helps
identify fragments of text which could plausibly be computed from data, including numerical values selected
from records or computed by aggregations like sum and mean, comparatives and superlatives like ``better than''
and ``largest'', trend-adjectives like ``growing'', and other idiomatic quantitative or semi-quantitative
phrases. An \InterpretationAgent, given such a fragment, then attempts to synthesise a suitable Fluid query
over the data which will generate the target string. The resulting expression is spliced into the source code
for an interactive web page, turning the static text fragment into an interactable data-driven element able to
reveal the data that underwrites the natural language claim. We evaluate our approach on a subset of SciGen,
an open source dataset consisting of tables from scientific articles and their corresponding descriptions,
which we extend with hand-generated counterfactual test cases to evaluate how well machine-generated
expressions generalise in the presence of changes to the data. Our results show that some state-of-the-art
models are often able to synthesise compound expressions that are extensionally compatible with our gold
solutions.
\end{abstract}