### 2025-09-18 paper/sec/appendix-prompts.tex
[update] change path of system prompts

\section*{System Prompt Interpretation Agent}

\VerbatimInput[fontsize=\small]{../system-prompt/interpretation-agent/system-prompt.txt}

\section*{System Prompt Recognition Agent}

\VerbatimInput[fontsize=\small]{../system-prompt/suggestion-agent/system-prompt.txt}