### 2025-09-18 paper/sec/appendix-prompts.tex

\section{\InterpretationAgent System Prompt}

\VerbatimInput[fontsize=\small]{../system-prompt/interpretation-agent/system-prompt.txt}

\section{\SuggestionAgent System Prompt}

\VerbatimInput[fontsize=\small]{../system-prompt/suggestion-agent/system-prompt.txt}