### 2025-09-19 paper/abstract.tex

\begin{abstract}
We introduce \emph{transparent documents}, interactive web-based scholarly articles which allow readers to
explore the relationship to the underlying data by hovering over fragments of text. We present an agent-based
LLM framework for authoring transparent documents, building on recent developments in data provenance for
general-purpose programming languages. Our implementation uses Fluid, an open source functional programming
language with a provenance-tracking runtime, as a target platform. Our tool consists of two LLM agents which
support a human author during the creation of a transparent document. A \SuggestionAgent helps identify
fragments of text which could plausibly be computed from data, including numerical values selected from
records or computed by aggregations like sum and mean, comparatives and superlatives like ``better than'' and
``largest'', trend-adjectives like ``growing'', and other idiomatic quantitative or semi-quantitative phrases.
An \InterpretationAgent, given such a fragment, then attempts to synthesise a suitable Fluid query over the
data which will generate the target string. The resulting expression is spliced into the source code for an
interactive web page, turning the static text fragment into an interactable data-driven element able to reveal
the data that underwrites the natural language claim. We evaluate our approach on a subset of SciGen, an open
source dataset consisting of tables from scientific articles and their corresponding descriptions, which we
extend with hand-generated counterfactual test cases to evaluate how well machine-generated expressions
generalise in the presence of changes to the data. Our results show that some state-of-the-art models are
often able to synthesise compound expressions extensionally compatible with our gold solutions.
\end{abstract}