### 2025-09-22 paper/sec/conclusion.tex
[new] section conclusion-further works with table

\section{Conclusion}
\label{sec:conclusion}

\input{fig/nl-further-works}