### 2025-09-23 paper/sec/authoring-workflow.tex
Tricky.

\section{AI-Assisted Authoring Workflow}
\label{sec:authoring-workflow}

Our authoring tool is composed of two LLM-based agents:

\begin{itemize}
    \item \textbf{\SuggestionAgent}. Identifies text fragments potentially computable from data.
    \item \textbf{\InterpretationAgent}. For a given text fragment (provided by SuggestionAgent or author),
    attempts to synthesise Fluid expression computing target fragment.
\end{itemize}

The main components of the workflow are as follows:

\begin{enumerate}

\item \textbf{Initial configuration.} Author imports text and accompanying data into the system, creating a
programmatic representation of the target document. Initially this representation is isomorphic to the target
text, namely a string literal of the form \kw{"""..."""}, where the triple quotes are Fluid syntax for an
\emph{interpolated string}, i.e.~a string literal where expressions of the form \kw{\{$e$\}} are permitted
within the string. The \SuggestionAgent analyses these inputs and identifies any fragments of natural language
which are candidates for being computed instead of remaining as literal substrings.

\item \textbf{High-level Authoring workflow.} The system then enters the human-in-the loop authoring workflow
shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
system initially waits for the author to select a fragment of text $s$ to interpret, a fragment which may or
may not have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
expression $e$ using a closed-loop synthesis step ((3) below). If this process succeeds, the system replaces
the selected substring $s$ with the interpolation expression \kw{\{$e$\}}, creating a new document
configuration. This is the primary successful path through the workflow and returns the system to the entry
state where it is waiting for another top-level interaction from the author.

\item \textbf{Code synthesis step.} The expression synthesis step is itself an internal loop which starts with
an initial prompt sent to the LLM (\secref{prompt-design} below) requesting the generation of an expression
$e$. Using the Fluid command-line interface, the expression is validated by checking that it evaluates without
error, produces a value coercible to a string $s'$, and finally that $s'$ is equal to the target fragment $s$.
Any failure triggers prompt augmentation with the appropriate error message and the system retries generation.
If code synthesis loop is able to yield an expression which computes $s$ within a maximum
number of retries, the system proceeds to the manual validation step (4). Otherwise, the code synthesis step has failed and there are two possibilities:
\begin{enumerate}
\item If the last generated $e$ was invalid (resulting in an error), no remedial action is open to the author. This is considered an unsuccessful path through the workflow and returns the system to the start state.
\item In the $s' \neq s$ case, the user can choose to manually abort and return to the entry state, or
optionally to \emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the
candidate expression. This is intended to cover the situation where the author has made a claim which is
\emph{incorrect}, but the data set and surrounding natural language has caused the LLM to synthesise an
expression which generates a different value from the one specified by the user.
\end{enumerate}

\item \textbf{Manual validation step.} Once a candidate expression has been generated, the author is able to
republish the web page containing the Fluid article and interact with the fragment of text. As shown in
\secref{}, this is an important validation step that can reveal errors in the generated expression.

\end{enumerate}

This human-in-the-loop design combines automated synthesis with validation and author oversight, providing a
substantial level of automation, but requiring the author to intervene at key steps to ensure correctness.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{fig/interpretation-agent}
    \caption{Human-in-the-Loop workflow (states requiring human intervention in grey)}
    \label{fig:interpretation-agent}
\end{figure}

\subsection{Prompt Design}
\label{sec:prompt-design}

\paragraph{System Prompt.}
For the Interpretation Assistant, the system prompt guides the LLM in replacing placeholder tags in the
supplied paragraph with executable Fluid expressions. It includes the following information about the document
under construction, the exact task to perform, and constraints on the expected output. The full system prompt
is given in \appref{system-prompt:interpretation-agent}.

\begin{itemize}
\item \textbf{Input structure:} summarises the imported datasets, modules, auxililiary definitions in scope in
the current file, the current Fluid representation of the target paragraph with \kw{[REPLACE …]} tag
indicating the text fragment to replace, plus the full text for the paragraph.
\item \textbf{Task description:} description of the LLM task, namely to identify \kw{[REPLACE …]} placeholder
and generate Fluid expression that produces target string using supplied dataset and definitions in scope.
\item \textbf{Output constraints:} specify that the LLM should return a syntactically valid Fluid expression,
with no extraneous content (such as comments).
\end{itemize}

\rpnote{Now something about the rest of the prompt.}

\begin{figure}[t]
    \centering
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-correct}
        \caption{Author Acceptance pathway}
        \label{fig:data-flow-correct}
    \end{subfigure}\hfill
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-error}
        \caption{Author Rejection pathway}
        \label{fig:dataflow-error}
    \end{subfigure}
    \caption{Two possible paths through the basic editing loop: (a) author accepts generated expression after
    interactively verifying its correctness; (b) author rejects proposed expression, having identified an
    error (natural language mentions accuracy, while interaction reveals use of timing data).}
    \label{fig:overall-architecture}
\end{figure}