### 2025-09-23 paper/sec/authoring-workflow.tex
wow this is hard.

\section{AI-Assisted Authoring Workflow}
\label{sec:authoring-workflow}

Our authoring tool is composed of two LLM-based agents:

\begin{itemize}
    \item \textbf{\SuggestionAgent}. Identifies text fragments potentially computable from data.
    \item \textbf{\InterpretationAgent}. For a given text fragment (provided by SuggestionAgent or author),
    attempts to synthesise Fluid expression computing target fragment.
\end{itemize}

The main components of the workflow are as follows:

\begin{enumerate}

\item \textbf{Initial configuration.} The author imports the target text and accompanying data into the
system, creating a programmatic representation of the target document. Initially this is simply isomorphic to
the target text, taking the form of a string literal \kw{"""..."""}, where the triple quotes are Fluid syntax
for a Python or JavaScript-style \emph{interpolated string}, i.e.~a literal where expressions of the form
\kw{\{$e$\}} are permitted within the string. The \SuggestionAgent analyses these inputs and identifies any
fragments of natural language which are candidates for being computed instead of remaining as literal
substrings.

\item \textbf{High-level Authoring workflow.} The system then enters the human-in-the loop authoring workflow
shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
system initially waits for the author to select a fragment of text $s$ to interpret (perhaps but not
necessarily highlighted by the \SuggestionAgent). The system then attempts to generates a candidate Fluid
expression $e$ using the closed-loop synthesis step (3) below. If this process succeeds with an expression
$e$, the system proceeds to the manual validation step (4) below. If the code generation step fails with no
expression, no remedial action is open to the author; this is considered an unsuccessful path through the
workflow and returns the system to the start state. Otherwise the code generation step fails with a mismatched
string $s' \neq s$ outcome, and the user can choose to manually abort and return to the entry state, or
optionally to \emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the
candidate expression. This is intended to cover the situation where the author has made a claim which is
\emph{incorrect}, but the data set and surrounding natural language has caused the LLM to synthesise an
expression which generates a different value from the one specified by the user.

\item \textbf{Code synthesis step.} The expression synthesis step is itself an internal loop which starts with
an initial prompt sent to the LLM (\secref{prompt-design} below) requesting the generation of an expression
$e$. Using the Fluid command-line interface, the expression is validated by checking that it evaluates without
error, produces a value coercible to a string $s'$, and finally that $s'$ is equal to the target fragment $s$.
Any failure triggers prompt augmentation with the appropriate error message and the system retries generation.
If code synthesis loop is able to yield an expression which computes $s$ within a maximum number of retries,
the synthesis step succeeds with $e$. If the last generated $e$ was invalid (resulting in an error), the code
synthesis step fails with no expression. Otherwise, the code synthesis fails with a mismatched string outcome.

\item \textbf{Manual validation step.} Once a candidate expression has been generated, the author is able to
republish the web page containing the Fluid article and interact with the fragment of text.  As shown in
\secref{}, this is an important validation step that can reveal errors in the generated expression. If the
interactions look reasonable, the author can approve the replacement of the selected substring $s$ with the
interpolation expression \kw{\{$e$\}}, creating a new document configuration. This is the primary successful
path through the workflow and returns the system to the entry state where it is waiting for another top-level
interaction from the author. Otherwise, the author can reject the proposed change and return to the start
state.

\end{enumerate}

This human-in-the-loop design combines automated synthesis with validation and author oversight, providing a
substantial level of automation, but requiring the author to intervene at key steps to ensure correctness.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{fig/interpretation-agent}
    \caption{Human-in-the-Loop workflow (states requiring human intervention in grey)}
    \label{fig:interpretation-agent}
\end{figure}

\subsection{Prompt Design}
\label{sec:prompt-design}

\paragraph{System Prompt.}
For the Interpretation Assistant, the system prompt guides the LLM in replacing placeholder tags in the
supplied paragraph with executable Fluid expressions. It includes the following information about the document
under construction, the exact task to perform, and constraints on the expected output. The full system prompt
is given in \appref{system-prompt:interpretation-agent}.

\begin{itemize}
\item \textbf{Input structure:} summarises the imported datasets, modules, auxililiary definitions in scope in
the current file, the current Fluid representation of the target paragraph with \kw{[REPLACE …]} tag
indicating the text fragment to replace, plus the full text for the paragraph.
\item \textbf{Task description:} description of the LLM task, namely to identify \kw{[REPLACE …]} placeholder
and generate Fluid expression that produces target string using supplied dataset and definitions in scope.
\item \textbf{Output constraints:} specify that the LLM should return a syntactically valid Fluid expression,
with no extraneous content (such as comments).
\end{itemize}

\rpnote{Now something about the rest of the prompt.}

\begin{figure}[t]
    \centering
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-correct}
        \caption{Author Acceptance pathway}
        \label{fig:data-flow-correct}
    \end{subfigure}\hfill
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-error}
        \caption{Author Rejection pathway}
        \label{fig:dataflow-error}
    \end{subfigure}
    \caption{Two possible paths through the basic editing loop: (a) author accepts generated expression after
    interactively verifying its correctness; (b) author rejects proposed expression, having identified an
    error (natural language mentions accuracy, while interaction reveals use of timing data).}
    \label{fig:overall-architecture}
\end{figure}