### 2025-09-24 paper/sec/conclusion.tex

\section{Conclusions and Future Work}
\label{sec:conclusion}

\subsection{Future Work}
\label{sec:conclusion:future-work}

Providing helper functions such as \kw{growShrink} and \kw{trendWord} as a predefined library is in some ways
beneficial because they offer a uniform framework for interpreting a given scholarly document. We would also
like make the system able to operate in the absence of these, perhaps adding a loopback condition to turn
``Definition not found'' errors into an augmented prompt to generate the appropriate definition.

Another important direction is to broaden the scope of the artifacts than can be referred to and the natural
language forms that can be used to refer to them. For example, many claims in scholarly writing refer to
visualisations and intermediate data sets derived via cleansing, imputation or aggregation. These should also
be made available to the LLM (for visualisations, as structured data rather than as static bitmaps), enabling
interpretation of natural language in terms of visual information and other intermediate analyses. Extending
coverage to additional idioms beyond those considered in Table~\ref{tab:natural-language-forms}, such as
cardinals and multiplicatives (\textfrag{three} and \textfrag{three times}), rounding (\textfrag{about 2
degrees warmer}) and graded adjectives \textfrag{extremely likely}, would also expand the expressivity of the
assistant.

Consider the false statement \textfrag{The population of France is the largest in the EU}. There are at least
two ways to repair this statement so that it comes out true: treat the subject (France) as fixed, but adjust
the predicate so that it matches the facts (\textfrag{second largest}), or treat the predicate as fixed, but
keep adjust the subject to match (\textfrag{Germany}). To make these sorts of (pragmatically important)
distinctions, we may need to introduce the idea of \emph{referential terms} that have fixed denotations, in
contrast to queries, whose values are contingent on data.

Finally, integration into developer and authoring environments such as VSCode or Cursor would make the
workflow more seamless for authors. Automatic generation of the counterfactual test cases discussed in
\secref{validation-with-counterfactual-data} could further strengthen validation at authoring-time, improving
the trustability of interpreted claims.