### 2025-09-24 paper/fig/scigen-1805.02474v1-10-src.tex

\begin{figure}%[h]
    \small
    % listings package seems to struggle, for now just use screenshot from VSCode
    \includegraphics[scale=0.169]{fig/scigen-1805.02474v1-10-src-screenshot.png}
    \caption{Gold solution for transparent text in \figref{scigen-example-website} (some lines omitted)}
    \label{fig:fluid-example-paragraph}
\end{figure}