### 2025-09-25 paper/sec/authoring-workflow.tex

\section{AI-Assisted Authoring Workflow}
\label{sec:authoring-workflow}

Our authoring tool is composed of two LLM-based agents:

\begin{itemize}
    \item \textbf{\SuggestionAgent}. Identifies text fragments potentially computable from data.
    \item \textbf{\InterpretationAgent}. For a given text fragment (provided by SuggestionAgent or author),
    attempts to synthesise Fluid expression computing target fragment.
\end{itemize}

The main components of the workflow are as follows:

\begin{enumerate}

\item \textbf{Initial configuration.} The author imports the target text and accompanying data into the system
to create a programmatic representation of the target document. Initially this is simply equivalent to the
target text, taking the form of a string literal \kw{"""..."""}, where the triple quotes are Fluid syntax for
a Python or JavaScript-style \emph{interpolated string}, i.e.~a literal where expressions of the form
\kw{\{$e$\}} are permitted within the string. The \SuggestionAgent analyses the target text and identifies any
fragments which are candidates for being computed instead of remaining as literal substrings.

\item \textbf{High-level Authoring workflow.} The system then enters the human-in-the loop authoring workflow
shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
system waits for the author to select a fragment of text $s$ to interpret (perhaps previously highlighted by
the \SuggestionAgent). The system then attempts to generate a candidate Fluid expression $e$ using the
closed-loop synthesis step (3) below. If code synthesis succeeds with an expression $e$, the system proceeds
to the manual validation step (4) below. If the synthesis step fails with no expression, no remedial action is
possible; this is considered an unsuccessful path through the workflow and returns the system to the entry
state. Otherwise the synthesis step produces an expression $e$ which evaluates to a mismatched string $s' \neq
s$ outcome, and the user can choose to manually abort and return to the entry state, or optionally to
\emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the candidate
expression. This is intended to cover the situation where the author has made a claim which is
\emph{incorrect}, and the data set and surrounding natural language have led the LLM to synthesise an
expression which generates a different value from the one specified by the user.

\item \textbf{Code synthesis step.} The expression synthesis step is an error-guided iterative prompting
loop~\citep{skreta23}, beginning with an initial prompt sent to the LLM (\secref{prompt-design} below)
requesting the generation of an expression $e$. Using the Fluid command-line interface, the expression is
validated to check that it evaluates without error, produces a value coercible to a string $s'$, and finally
that $s'$ is equal to the target fragment $s$. Any failure triggers prompt augmentation with the appropriate
error message and the system retries generation. If code synthesis loop is able to yield an expression which
computes $s$ within a maximum number of retries, the synthesis step succeeds with $e$. If the last generated
$e$ was invalid (resulting in an error), the code synthesis step fails with no expression. Otherwise, code
synthesis produces an expression $e$ but with a mismatched string outcome $s' \neq s$.

\item \textbf{Manual validation step.} Once a candidate expression has been generated, the system replaces the
selected substring $s$ with the interpolation expression \kw{\{$e$\}}, creating a new (but only tentative)
document configuration. The author can republish the web page hosting the document and interact with the
proposed revision. As shown in \secref{}, this is an important validation step that can reveal errors in the
generated expression. If the interactions look reasonable, the author can approve the new document state; this
is the primary successful path through the workflow and returns the system to the entry state where it is
waiting for another top-level interaction from the author. Otherwise, the author rejects the proposed change
and returns to the entry state without any change to the document.

\end{enumerate}

This human-in-the-loop design combines automated synthesis with validation and author oversight, providing a
substantial level of automation, but requiring the author to intervene at key steps to ensure correctness.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{fig/interpretation-agent}
    \caption{Human-in-the-Loop workflow (states requiring human intervention in grey)}
    \label{fig:interpretation-agent}
\end{figure}

\subsection{The InterpertationAgent}
\label{sec:interpretation-agent}

\cdnote{I integrated the prompt description in the description of the InterpretationAgentv (unsure if this is the best plece) -- the prompt description seemed to basic (it's commented if you want to revert)}

The InterpretationAgent is guided by a structured system prompt that
frames code generation as a precise replacement task. The model
receives the imported datasets, helper modules, and the current Fluid
representation of the paragraph, in which a text fragment is marked
with [REPLACE …]. Its task is to substitute this placeholder with a
Fluid expression that evaluates exactly to the target string,
reconstructing quantitative or comparative claims as dataset queries
where appropriate. To ensure integration with the workflow, the output
must consist solely of a syntactically valid Fluid expression, with no
additional commentary. This focused design constrains the model while
providing enough context to generate expressions that are both
data-grounded and immediately executable; the full prompt is given in~\appref{system-prompt:interpretation-agent}.

%\paragraph{System Prompt.}
%For the Interpretation Assistant, the system prompt guides the LLM in replacing placeholder tags in the
%supplied paragraph with executable Fluid expressions. It includes the following information about the document
%under construction, the exact task to perform, and constraints on the expected output. The full system prompt
%is given in \appref{system-prompt:interpretation-agent}.

%\begin{itemize}
%\item \textbf{Input structure:} summarises the imported datasets, modules, auxililiary definitions in scope in
%the current file, the current Fluid representation of the target paragraph with \kw{[REPLACE …]} tag
%indicating the text fragment to replace, plus the full text for the paragraph.
%\item \textbf{Task description:} description of the LLM task, namely to identify \kw{[REPLACE …]} placeholder
%and generate Fluid expression that produces target string using supplied dataset and definitions in scope.
%\item \textbf{Output constraints:} specify that the LLM should return a syntactically valid Fluid expression,
%with no extraneous content (such as comments).
%\end{itemize}

%\rpnote{Now something about the rest of the prompt.}

\begin{figure}[t]
    \centering
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-correct}
        \caption{Author Acceptance pathway}
        \label{fig:data-flow-correct}
    \end{subfigure}\hfill
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-error}
        \caption{Author Rejection pathway}
        \label{fig:dataflow-error}
    \end{subfigure}
    \caption{Two possible paths through the basic editing loop: (a) author accepts generated expression after
    interactively verifying its correctness; (b) author rejects proposed expression, having identified an
    error (natural language mentions accuracy, while interaction reveals use of timing data).}
    \label{fig:overall-architecture}
\end{figure}