### 2025-09-25 paper/sec/authoring-workflow.tex

\section{AI-Assisted Authoring Workflow}
\label{sec:authoring-workflow}

Our authoring tool is composed of two LLM-based agents. A \textbf{\SuggestionAgent} identifies text fragments
potentially computable from data, and an \textbf{\InterpretationAgent}, given a text fragment provided by the
SuggestionAgent or by the author, attempts to synthesise a Fluid expression which computes the target
fragment. The main components of the workflow are as follows:

\begin{enumerate}

\item \textbf{Initial configuration.} The author imports the target text and accompanying data into the system
to create a programmatic representation of the target document. Initially this is simply equivalent to the
target text, taking the form of a string literal \kw{"""..."""}, where the triple quotes are Fluid syntax for
a Python or JavaScript-style \emph{interpolated string}, i.e.~a literal where expressions of the form
\kw{\{$e$\}} are permitted within the string. The \SuggestionAgent analyses the target text and identifies any
fragments which are candidates for being computed instead of remaining as literal substrings.

\item \textbf{High-level Authoring workflow.} The system then enters the human-in-the loop authoring workflow
shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
system waits for the author to select a fragment of text $s$ to interpret (perhaps previously highlighted by
the \SuggestionAgent). The system then attempts to generate a candidate Fluid expression $e$ using the
closed-loop synthesis step (3) below. If code synthesis succeeds with an expression $e$, the system proceeds
to the manual validation step (4) below. If the synthesis step fails with no expression, no remedial action is
possible; this is considered an unsuccessful path through the workflow and returns the system to the entry
state. Otherwise the synthesis step produces an expression $e$ which evaluates to a mismatched string $s' \neq
s$ outcome, and the user can choose to manually abort and return to the entry state, or optionally to
\emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the candidate
expression. This is intended to cover the situation where the author has made a claim which is
\emph{incorrect}, and the data set and surrounding natural language have led the LLM to synthesise an
expression which generates a different value from the one specified by the user.

\item \textbf{Code synthesis step.} The expression synthesis step is an error-guided iterative prompting
loop~\citep{skreta23}, beginning with an initial prompt sent to the LLM (\secref{prompt-design} below)
requesting the generation of an expression $e$. Using the Fluid command-line interface, the expression is
validated to check that it evaluates without error, produces a value coercible to a string $s'$, and finally
that $s'$ is equal to the target fragment $s$. Any failure triggers prompt augmentation with the appropriate
error message and the system retries generation. If code synthesis loop is able to yield an expression which
computes $s$ within a maximum number of retries, the synthesis step succeeds with $e$. If the last generated
$e$ was invalid (resulting in an error), the code synthesis step fails with no expression. Otherwise, code
synthesis produces an expression $e$ but with a mismatched string outcome $s' \neq s$.

\item \textbf{Manual validation step.} Once a candidate expression has been generated, the system replaces the
selected substring $s$ with the interpolation expression \kw{\{$e$\}}, creating a new (but only tentative)
document configuration. The author can republish the web page hosting the document and interact with the
proposed revision. As shown in \secref{}, this is an important validation step that can reveal errors in the
generated expression. If the interactions look reasonable, the author can approve the new document state; this
is the primary successful path through the workflow and returns the system to the entry state where it is
waiting for another top-level interaction from the author. Otherwise, the author rejects the proposed change
and returns to the entry state without any change to the document.

\end{enumerate}

This human-in-the-loop design combines automated synthesis with validation and author oversight, providing a
substantial level of automation, but requiring the author to intervene at key steps to ensure correctness.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{fig/interpretation-agent}
    \caption{Human-in-the-Loop workflow (states requiring human intervention in grey)}
    \label{fig:interpretation-agent}
\end{figure}

\paragraph{\InterpretationAgent prompt design.}

The \InterpretationAgent is guided by a structured system prompt that frames code generation as a precise
replacement task. The model receives the imported datasets, helper modules, and the current Fluid
representation of the paragraph, in which a text fragment is marked with the tag \kw{[REPLACE …]}. Its task is
to substitute this placeholder with a Fluid expression that evaluates exactly to the target string,
reconstructing quantitative or comparative claims as data queries. To ensure integration with the workflow,
the output must consist solely of a syntactically valid Fluid expression, with no additional commentary. The
full prompt is given in~\appref{system-prompt:interpretation-agent}.

%\paragraph{System Prompt.}
%For the Interpretation Assistant, the system prompt guides the LLM in replacing placeholder tags in the
%supplied paragraph with executable Fluid expressions. It includes the following information about the document
%under construction, the exact task to perform, and constraints on the expected output. The full system prompt
%is given in \appref{system-prompt:interpretation-agent}.

%\begin{itemize}
%\item \textbf{Input structure:} summarises the imported datasets, modules, auxililiary definitions in scope in
%the current file, the current Fluid representation of the target paragraph with \kw{[REPLACE …]} tag
%indicating the text fragment to replace, plus the full text for the paragraph.
%\item \textbf{Task description:} description of the LLM task, namely to identify \kw{[REPLACE …]} placeholder
%and generate Fluid expression that produces target string using supplied dataset and definitions in scope.
%\item \textbf{Output constraints:} specify that the LLM should return a syntactically valid Fluid expression,
%with no extraneous content (such as comments).
%\end{itemize}

%\rpnote{Now something about the rest of the prompt.}

\begin{figure}[t]
    \centering
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-correct}
        \caption{Author Acceptance pathway}
        \label{fig:data-flow-correct}
    \end{subfigure}\hfill
    \begin{subfigure}{0.50\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/data-flow-error}
        \caption{Author Rejection pathway}
        \label{fig:dataflow-error}
    \end{subfigure}
    \caption{Two possible paths through the basic editing loop: (a) author accepts generated expression after
    interactively verifying its correctness; (b) author rejects proposed expression, having identified an
    error (natural language mentions accuracy, while interaction reveals use of timing data).}
    \label{fig:overall-architecture}
\end{figure}