### 2025-09-25 paper/sec/related-work.tex
Related work tweaks.

\section{Related Work}
\label{sec:related}

\paragraph{Argument mining.}
\emph{Argument mining} is an area of NLP which involves identifying argumentative structures in text, such as
claims, premises, and conclusions, and mapping them to formal
representations~\citep{palau2009argumentation,lippi2015argument}. Early work focused on rule-based approaches,
while more recent work has leveraged machine learning and deep learning
techniques~\citep{stab2014identifying,stab2017parsing,eger2017neural}. The field has also emphasized defining
annotation schemes for the task, such as the Argumentative Zoning
framework~\citep{teufel2002summarizing,teufel2009towards}, as well as schemes more directly tailored to
argument mining~\citep{stab2014annotating}. The field has focused on various domains, starting from legal
texts~\citep{toulmin2003uses}, and has relied on online resources such as
Debatepedia~\citep{cabrio2013natural}. The community has also rapidly engaged with work that explores the use
of argument mining in scientific texts~\citep{liakata12,lauscher2018investigating} to better understand the
structure of scientific arguments and the relationships between different claims and evidence. While the
advent of LLMs has improved performance~\citep{gorur2025can,vrakatseli2025can}, argument mining remains a
challenging task, particularly when it comes to identifying implicit argumentative relations between discourse
units, and reasoning about relationships among different argumentative components, especially in cross-domain
settings where models struggle to generalise~\citep{gemechu2024aries}. While in our work we do not directly
perform traditional argument mining, this work similarly relies on the identification of claims in text, which
we evaluate following established practices in the field.

\paragraph{NLP and scientific writing.}
The intersection of NLP and scientific writing has gained increasing attention in the last decade, with a
focus on improving the clarity, coherence, and overall quality of scientific texts. On the authoring side,
tools such as automated writing assistants can support researchers in producing more fluent and accessible
text, for instance through grammar correction, summarisation and text
simplification~\citep{napoles2017jfleg,stiennon2020learning,takeshita2024cross,saggion2017automatic}. Other
approaches specifically target the argumentative structure of scientific papers, helping writers to organise
contributions and claims more effectively~\citep{lauscher2018arguminsci}. Nowadays, general purpose LLMs such
as ChatGPT or tools tailored for the task such as Grammarly show the variety of support that NLP tools can
provide to authors~\citep{wu2023chatgpt,ahn2024transformative,khalifa2024using}.

At the same time, NLP methods are being developed to assist reviewers and editors in evaluating submissions.
These include systems for detecting potential issues such as lack of clarity, weak argumentative support, or
even factual inconsistencies and up to scientific fraud~\citep{thakkar2025can, fromm2021argument,
freedman2024detecting}. Such tools can also facilitate meta-reviewing by providing summaries of peer reviews
and identifying points of disagreement among reviewers~\citep{kumar2023reviewers}. While AI tools show
promises in improving the peer-review process~\citep{tyser2024ai}, there are also various risks associated
such as breaches of confidentiality, lack of transparency and biases~\citep{perlis2025artificial}. Our current
work situates itself in the context of NLP tools for supporting the understanding of scientific writing;
specifically, it addresses one of the major critiques toward the automation of such process by offering a
transparent way of examining its workflow.

\paragraph{Interpretable NLP.}

As \figref{scigen-example-website} illustrates, scientific texts routinely make use of comparatives like
``faster'' while leaving one of the argument slots implicit, with the context determining the omitted
referent. LLMs demonstrate considerable competence in resolving these and other more syntactic forms of
anaphora such as pronouns~\citep{zhu25}, but the resolved referent itself -- concretely, what was being
referred to -- remains implicit. Interpretable NLP is a recent research direction which aims to support
comprehension (and production) of text in a more explicit and transparent way~\citep{yulan23}. By generating
code that formalises the interpretation of a comparative like ``faster'', our approach also makes these
implicit references explicit; combining our system with interpretable NLP would allow the user to explore the
linguistic interpretation as well.