### 2025-09-25 paper/sec/evaluation.tex

\section{Experimental Evaluation}
\label{sec:evaluation}

\subsection{Research questions}

We investigated the following research questions, which probe the limits of the \InterpretationAgent's ability
to reliably provide a data-driven interpretation of natural language. By contrasting genuine uncertainty with
misleading or ill-specified inputs, we aim to understand not only the model's raw success rate, but also the
failure modes most likely to arise in practical use.

\paragraph{RQ1. How robust is the system when provided with incomplete or ambiguous information?}

In real-world authoring tasks, text fragments may omit key criteria (e.g. \textbf{?}) or contain vague
identifiers (e.g. \textbf{?} without further context). We therefore ask how performance degrades when the
specification is underspecified or when multiple interpretations are possible. Here, we distinguish between
two forms of ambiguity: (1) domain-level ambiguity, which reflects genuine uncertainty in the data (e.g. two
models achieving identical scores). (2) specification-level ambiguity, which arises from incomplete or
imprecise descriptions in the query (e.g. missing fields that would otherwise resolve the expression
uniquely).

\paragraph{RQ2. How sensitive is the system to misleading or adversarial information?}

Authors may (intentionally or unintentionally) introduce misleading elements into the text. These include
inconsistent use of language, incorrect references to numerical values, or contradictory claims. We ask how
the system responds in such cases: does it fail gracefully, does it default to literal reproduction of the
misleading fragment, or does it attempt to generate a plausible, but incorrect, expression?
Table~\ref{tab:problematic_cases} shows some of cases which are ``problematic'' in this sense; in these cases,
producing a valid expression is likely to be challenging because of ambiguities in the input data or
accompanying natural language.

\begin{figure}
    \centering
    \begin{subfigure}{0.48\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/success_rate_by_category}
        \caption{Success rate by Linguistic Category}
        \label{fig:success_rate_by_category}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\linewidth}
        \centering
        \includegraphics[width=\linewidth]{fig/success_rate_by_complexity}
        \caption{Success rate by Complexity}
        \label{fig:success_rate_by_complexity}
    \end{subfigure}
    \caption{Success rate of the proposed system, measured over 5 runs with \gptfour}
    \label{fig:success_rate_comparison}
\end{figure}

\subsection{Results}

%\begin{table}[h!]
%    \scriptsize
%
%    \centering
%    \begin{subtable}{0.45\linewidth}
%        \centering
%        \begin{tabular}{lcccc}
%            \hline
%            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
%            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
%            \hline
%            Comparison     & 68.2   & 3.2 & 77.4 & 2 \\
%            Data Retrieval & 77.3   & 3.7 & 77.4 & 2 \\
%            Difference     & 20.0   & 44.7 & 77.4 & 2 \\
%            Min/max        & 97.0   & 5.6 & 77.4 & 2 \\
%            Rank           & 0.0   & 0.0 & 77.4 & 2 \\
%            Sum            & 100   & 0.0 & 77.4 & 2 \\
%            Ratio          & 100   & 0.0 & 77.4 & 2 \\
%            Agerage        & 100   & 0.0 & 77.4 & 2 \\
%            Overall        & 74.9 & 3 & 77.4 & 2 \\
%            \hline
%        \end{tabular}
%        \caption{Results with target-value = true}
%        \label{tab:results_true}
%    \end{subtable}%
%    \hfill
%    \begin{subtable}{0.45\linewidth}
%        \centering
%        \begin{tabular}{lcccc}
%            \hline
%            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
%            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
%            \hline
%            Comparison     & 33.3 & 3.1 & -- & -- \\
%            Data Retrieval & 62.0 & 1.6 & -- & -- \\
%            Difference     & 75.0 & 50.0 & -- & -- \\
%            Min/max        & 90.6 & 12.6 & -- & -- \\
%            Rank           & 0.0 & 0.0 & -- & -- \\
%            Sum            & 100.0 & 0.0 & -- & -- \\
%            Ratio          & 100.0 & 0.0 & -- & -- \\
%            Average        & 100.0 & 0.0 & -- & -- \\
%            Overall        & 57.1 & 0.8 & -- & -- \\
%            \hline
%        \end{tabular}
%        \caption{Results without target-value}
%        \label{tab:results_no_target}
%    \end{subtable}
%    \caption{Success rates (with standard deviation) for \textbf{\gptfour} and \textbf{\gptfive} across different task categories}
%    \label{tab:results_combined}
%\end{table}

To evaluate the system we used a sample of the \SciGen dataset~\citep{moosavi21}. We aggregated the generated
expression in the linguistic categories from Table~\ref{tab:natural-language-forms}.
\figref{success_rate_by_category} shows the results obtained in generating expression by categories.

%Table \ref{tab:results_combined} summarizes the performance of \gptfour and the most recent OpenAI model, \gptfive, under two different settings: with target-value sharing enabled and without target-value sharing.
%Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%).

%Figure~\ref{fig:success_rate_by_category} illustrates the success rate of \gptfour % under two different settings: with target-value sharing enabled and without target-value sharing.
%for each category, both with and without target-value sharing.
%Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%) \cdnote{this needs to be updated for target sharing and not target sharing or removed}.

%Table \ref{tab:results_combined} summarizes the performance of \gptfour and the most recent OpenAI model,
%\gptfive, under two different settings: with target-value sharing enabled and without target-value sharing.
%Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%).

In this section, we report the experimental results.
Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and without target-value sharing.
The results show that providing the target value is crucial for achieving a high success rate.
Moreover, the \InterpretationAgent generates a correct expression in approximately XX\% of the cases.

The system achieved performance above 70\% in comparison, data retrieval, and min/max search tasks.
In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (50\%).

Figure~\ref{fig:success_rate_by_complexity} reports the success rate as a function of the number of categories assigned to each expression.
Fluid instructions combining multiple operations were associated with multiple categories.
Results show that the success rate was 75\% for simple expressions (one category), increased to 85\% for two categories,
and dropped to 58\% when three categories were involved.

In addition to evaluating performance when the target expression is provided, we also assessed the model’s ability to produce a
correct expression when no reference string was given.
On the \SciGen dataset, the model achieved a 74.9\% success rate when provided with the target string.
Without the target string, performance decreased moderately to 39.4\%, indicating that the model retains substantial
ability to reconstruct valid solutions even without explicit guidance

\subsection{Counterfactual testing}
\label{sec:validation-with-counterfactual-data}

For the \emph{counterfactual testing} part of our analysis, the input data is deliberately modified in order
to verify the consistency of the behaviours. Both the expected and generated expressions are executed again
and the results are compared. If discrepancies are found the \InterpretationAgent ask for a new generation
sending back that the failure is related to a counterfactual error.

%\rpnote{complete}

Across 300 test executions, 121 contained at least one counterfactual error (avg. 3.8), of which 42 ultimately succeeded.
Such tests help detect expressions that may yield the expected output but not the correct solution.
For example, the \texttt{\textbackslash InterpretationAgent} generated:

\begin{verbatim}
(findWithKey_ "model" "LSTM2" tableData).time_s
\end{verbatim}

intended to retrieve the execution time of the \texttt{LSTM} model, but incorrectly referred to \texttt{LSTM2}.
Counterfactual testing exposed this error and prompted regeneration.

\begin{table}[t]
    \centering
    \small
    \begin{tabular}{l p{6cm} p{4cm}}
        \hline
        \textbf{Problem Type} & \textbf{Example} & \textbf{Explanation} \\
        \hline

        ambiguous referent &
        As shown in Table 3 \hl{BiLSTM} gives significantly better accuracies compared to uni-directional LSTM with the training time per epoch growing from 99 seconds to 106 seconds &
        There are three models that have 99 seconds as their time; the assistant would not know which one to choose. \\

        missing data &
        The average methane emissions for the year 2015 is \hl{1.0} &
        November is missing to calculate the average correctly. \\

        false statement &
        We additionally make comparisons with stacked CNNs and hierarchical attention (Vaswani et al., 2017), shown in Table 3 (the CNN and Transformer rows), BiLSTM is \hl{the most efficient} among all models compared, with the highest model size &
        BiLSTM is not the most efficient, nor does it have the largest size. \\

        incorrect numerical value &
        LSTM is the fastest model with overall time taken being \hl{90} seconds &
        It is not 90 but 106. \\

        ambiguous referent &
        LSTM is the fastest model with overall time taken being \hl{90} seconds &
        There are two type of time in the dataset (training\_time, execution\_time), both with a value of 90 seconds. \\

        \hline
    \end{tabular}
    \caption{Categories of problematic example}
    \label{tab:problematic_cases}
\end{table}