### 2025-09-25 paper/fig/natural-language-forms.tex

\begin{table}[!ht]
    \centering
    \footnotesize
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{>{\raggedright\arraybackslash}p{1.4cm} >{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{6.3cm}}
        \hline
        \textbf{Label}                & \textbf{Example} & \textbf{Gold Solution for Example} \\
        \rowcolor{gray!20}
%        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Quantitative expressions}} \\
        Data retrieval
        & the training time per epoch growing from \hl{67} seconds to 106 seconds.
        &
        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
(model_ "LSTM").time_s
        \end{lstlisting}
        \\
        Ratio &
        The Energy Sector accounts for total methane emissions of \hl{52.80}\% in 2030.
        &
        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
(getByCategory "Energy Sector" year).emissions /
  sum (map (fun x -> x.emissions)
    (getByYear year)) * 100

        \end{lstlisting}  \\
        \rowcolor{gray!20}
%        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Aggregation}} \\
        Average
        & The average methane emissions for the year 2030 is \hl{13.51} &
        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
sum (map (fun x -> x.emissions)
  (getByYear year)) / length records
        \end{lstlisting} \\
        Min/Max                          & The Energy Sector recorded its highest methane emissions in \hl{2030}             &
        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
let maxEntry = maximumBy (fun x -> x.emissions)
   (filter (fun x -> x.type == "Energy Sector")
      tableData)
in maxEntry.year
        \end{lstlisting} \\
        \rowcolor{gray!20}
        Rank &
        3-layer stacked CNN gives an accuracy of 81.46\%, which is the \hl{lowest} compared with BiLSTM, and S-LSTM  &
        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
rankLabel "lowest"
  (findIndex "model" "CNN"
    (sort cmpTime tableData)) \end{lstlisting} \\
        Sum &
        The total methane emissions for the year 2030 is \hl{37.74} for Agriculture &
        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
sum (map (fun x -> x.emissions)
   (getByYear year))
        \end{lstlisting} \\
        \rowcolor{gray!20}
%        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Trends}} \\
        Comparison
        & The training time per epoch \hl{growing} from 67 seconds to 106 seconds. &
        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
trendWord
  (model_ "BiLSTM" tableData).time_s
  (model_ "LSTM" tableData).time_s
  growShrink
        \end{lstlisting} \\
        Generalised quantifiers
        & In the case of one syndrome (Hemorrhagic) we noticed an \hl{unusually low} level of recall for SVM but not for NB. &
        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
unusuallyHighLow (overallComparison [
  compareCols col "naive_bayes_r"
    (findWithKey_ "synd" "Hem." tableData)
  | col <- ["svm1_r", "svm2_r", "svm3_r", "svmr_r"]
])
        \end{lstlisting} \\
        \hline
    \end{tabular}
    \caption{Quantitative/semi-quantitative natural language forms considered in this paper}
    \label{tab:natural-language-forms}
\end{table}